Under-approximation of Reachability in Multivalued Asynchronous Networks

### 2014-03-17 parts/models.tex
continue flattening ; only the big proof left

In \pref{def:fop}, we finally define the the operator $\Fopsymbol$ which characterizes the playability of an action
given the semantics of PH (\pref{def:play}).
\begin{definition}[Playability property of an action ($\Fopsymbol : \PHh \rightarrow \F$)]\label{def:fop}
  $$\forall h \in \ov{\PHh}, \Fop{h} \equiv \hitter{h} \wedge
    \left( \bigwedge_{\substack{g \in \PHh^{(n)}\\n < \prio(h)}}
    \neg \left( \target{g} \wedge \hitter{g}\right) \right)$$
\end{definition}

By definition of the semantics of a PH given in \pref{def:play},
an action $h \in \ov{\PHh}$ is playable in a state $s \in \ov{\PHl}$ iff $\Fsem{\Fop{h}}{s}$.

@@

Because we only use classical logic operators, we can compute the PP on any action in $\ov{\PH}$.
For any action $h \in \ov{\PHh}$, this DNF takes the form:
\[\Fop{h} \equiv \bigvee_{i \in \segm{1}{\n}} \left( \bigwedge_{j \in \segm{1}{\m}} p_{i,j} \right)\]
where $\n \in \sN$ and $\forall i \in \segm{1}{\n}, \m \in \sN^*$.
If $\n = 0$, then $\Fop{h} \equiv \bot$; this means that $h$ can never be played
due to preemptions by other actions with higher priorities.
If $\Fop{h} \not\equiv \bot$, on the other hand, then in this case $\Fop{h}$
can be seen as a disjunction of $\n$ smaller playability properties consisting only of conjunctions of atoms.
These $\n$ conjunctions can be translated to as many prioritized cooperative sorts,
thus creating a new PH model which is entailed in the restrictions depicted in \pref{ssec:hypothesis}.
%Indeed, for any $i \in \segm{1}{\n}$, the PP $\bigwedge_{j \in \segm{1}{\m}} p_{i,j}$
In this second case, we denote, for any $i \in \segm{1}{\n}$:
$\PHdep{i}{h} = \{ \PHsort(p_{i,j}) \mid j \in \segm{1}{\m} \}$.

\begin{definition}
  \label{def:flattening}
  If $k \in \sN^*$, we note $\PHflat(\ov{\PHs}; \ov{\PHl}; \ov{\PHa}^{\angles{k}}) = (\PHs; \PHl; \PHa^{\angles{2}})$
  the flattening of the PH with $k$ classes of priorities $\ov{\PHs}; \ov{\PHl}; \ov{\PHa}^{\angles{k}}$, with:
  \begin{itemize}
    \item $\PHs = \ov{\PHs} \cup \{ f^{h,i} \mid h \in \ov{\PHh} \wedge \n \geq 1 \wedge i \in \segm{1}{\n} \}$;
    \item $\PHl = \underset{a \in \PHs}{\times} \PHl_{a}$, where
      $\PHl_{a} = \ov{\PHl}_{a}$ if $a \in \ov{\PHs}$; and 
      $\PHl_{f^{h,i}} = \{ f^{h,i}_\varsigma \mid \varsigma \in \underset{b \in \PHdep{i}{h}}{\times} \PHl_{b} \}$ otherwise.
    \item $\PHh^{(1)} = \{ \PHhit{a_k}{f^{h,i}_\varsigma}{f^{h,i}_{\varsigma'}} \mid
      h \in \ov{\PHh} \wedge %\n \geq 1 \wedge i \in \segm{1}{\n} \wedge
      a \in \PHdep{i}{h} \wedge a_k \in \PHl_a \wedge
      f^{h,i}_\varsigma , f^{h,i}_{\varsigma'} \in \PHl_{f^{h,i}} \wedge
      \PHget{\varsigma}{a} \neq a_k \wedge \varsigma' = \varsigma \Cap \{ a_k \} \}$;
%      \todo{From here: ...} \PHget{\varsigma'}{a} = a_k \wedge
%      (\forall b \in \PHdep{i}{h} \setminus \{ a \}, \PHget{\varsigma'}{b} = \PHget{\varsigma}{b}) \}$;
%      \todo{... can be replaced by: $\varsigma' = \varsigma \Cap \{ a_k \}$}
    \item $\PHh^{(2)}=\{ \PHhit{f^{h,i}_\varsigma}{\target{h}}{\bounce{h}} \mid
      h \in \ov{\PHh} \wedge f^{h,i}_\varsigma \in \PHl_{f^{h,i}} \wedge \Fsem{h}{\varsigma} \}$.

    \item $\Hits^{(1)}= \{ \hit{a^j_k}{f^i_\varsigma}{f^i_{\varsigma'}}
      \mid i\in\segm{1}{n} \wedge
      j\in\DNdep(f^i) \wedge a^j_k\in \PHl_{a^j} \wedge f^i_\varsigma\in \PHl_{f^i}
      \wedge
      \get{\varsigma}{a^j}\neq a^j_k 
      \wedge \get{\varsigma'}{a^j}=a^j_k
      \wedge (\get{\varsigma'}{a^{l}}=\get{\varsigma}{a^l},
      \forall l\in\segm{1}{n},l\neq j)
      \}$, 
      the set of actions with priority $1$ that update cooperative sorts;

          \item $\Hits^{(2)}=\{ \hit{f^i_\varsigma}{a^i_k}{a^i_{k'}} \mid i\in\segm{1}{n} \wedge
          f^i_\varsigma\in \PHl_{f^i} \wedge
          a^i_k\in \PHl_{a^i}\wedge k\neq k' \wedge f^i(\decode \varsigma) = k'
        \}$,

  \end{itemize}

\end{definition}

@@

\begin{comment}

%\newcommand{\poz}{\F}
%\newcommand{\pon}{P_{DNF}}
%\newcommand{\unneg}{\mathsf{UnNeg}}
%\newcommand{\negate}{\mathsf{Negate}}
%\newcommand{\prednf}{\mathsf{Distribute}}
%\newcommand{\dnf}{\mathsf{DNF}}
%\newcommand{\DNF}{\dnf}
%\newcommand{\dnff}{\delta\F}

%\newcommand{\fun}{\mathcal{F}}

\begin{definition}[Equivalent Process Hitting]\label{def:equiv}
  Two PH models $\PH = (\PHs; \PHl; \PHh^\angles{k})$ and $\ov{\PH} = (\PHs; \PHl; \ov{\PHh}^\angles{n})$,
  where $k, n \in \sN^*$,
  are \emph{equivalent} iff:
  $$\forall s \in \PHl, ( \exists h = \PHhit{A}{b_i}{b_j} \in \PHh, \Fsem{\F(h)}{s}
    \Longleftrightarrow \exists g = \PHhit{\ov{A}}{b_i}{b_j} \in \ov{\PHh}, \Fsem{\F(g)}{s} )$$
\end{definition}

\begin{demo}
  Let $\PH = (\PHs; \PHl; \PHh^\angles{k})$ and $\ov{\PH} = (\PHs; \PHl; \ov{\PHh}^\angles{n})$
  two equivalent PH models, with $k, n \in \sN^*$.
  We want to show:
  $$\forall s, s' \in \PHl, s \neq s', s \trans{\PH} s' \Longleftrightarrow s \trans{\ov{\PH}} s' $$
  Let $s, s' \in \PHl$ with $s \neq s'$.
  ($\Rightarrow$) Suppose $s \trans{\PH} s'$.
  It means that there exists $h = \PHhit{A}{b_i}{b_j} \in \PHh$ so that $s' = s \PHplay h = s \Cap b_j$
  in $\PH$ (\pref{def:play}), \ie $\Fsem{\F(h)}{s}$ and $b_i \in s$.
  As $\ov{\PH}$ is equivalent to $\PH$,
  there exists $g = \PHhit{\ov{A}}{b_i}{b_j} \in \ov{\PHh}$ so that $\Fsem{\F(g)}{s}$ (\pref{def:equiv}).
  Therefore, $g$ is playable in $s$ and $s \PHplay g = s \Cap b_j = s'$;
  which means: $s \trans{\ov{\PH}} s'$
  ($\Leftarrow$) Same reasoning.
\end{demo}

\end{comment}

>>>>

In \pref{def:fop}, we finally define the the operator $\Fopsymbol$ which characterizes the playability of an action
given the semantics of PH (\pref{def:play}).
\begin{definition}[Playability property of an action ($\Fopsymbol : \PHh \rightarrow \F$)]\label{def:fop}
  $$\forall h \in \ov{\PHh}, \Fop{h} \equiv \hitter{h} \wedge
    \left( \bigwedge_{\substack{g \in \PHh^{(n)}\\n < \prio(h)}}
    \neg \left( \target{g} \wedge \hitter{g}\right) \right)$$
\end{definition}
%
\begin{theorem}
\label{th:ppplay}
  $h \in \ov{\PHh}$ is playable in $s \in \ov{\PHl}$ if and only if $\Fsem{\Fop{h}}{s}$.
\end{theorem}
%
\begin{proof}
  By definition of the semantics of a PH given in \pref{def:play}.
\end{proof}

@@

Because we only use classical logic operators, we can compute the Disjunctive Normal Form (DNF) of any PP.
For any action $h \in \ov{\PHh}$, this DNF takes the form:
\[\Fop{h} \equiv \bigvee_{i \in \segm{1}{\n}} \left( \bigwedge_{j \in \segm{1}{\m}} p_{i,j} \right)\]
where $\n \in \sN$ and $\forall i \in \segm{1}{\n}, \m \in \sN^*$.
If $\n = 0$, then $\Fop{h} \equiv \bot$; this means that $h$ can never be played
due to preemptions by other actions with higher priorities.
If $\Fop{h} \not\equiv \bot$, on the other hand, then in this case $\Fop{h}$
can be seen as a disjunction of $\n$ smaller playability properties consisting only of conjunctions of atoms.
These $\n$ conjunctions can be translated to as many prioritized cooperative sorts,
thus creating a new PH model which is entailed in the restrictions depicted in \pref{ssec:hypothesis}.
%Indeed, for any $i \in \segm{1}{\n}$, the PP $\bigwedge_{j \in \segm{1}{\m}} p_{i,j}$
In this second case, we denote, for any $i \in \segm{1}{\n}$:
$\PHdep{i}{h} = \{ \PHsort(p_{i,j}) \mid j \in \segm{1}{\m} \}$.

\newcommand{\flats}[1]{\lfloor #1 \rfloor}
\newcommand{\unflats}[1]{\lceil #1 \rceil}

\newcommand{\os}{\ov{s}}
\newcommand{\oPH}{\ov{\PH}}

\begin{definition}
  \label{def:flattening}
  If $k \in \sN^*$, we note $\PHflat(\ov{\PHs}; \ov{\PHl}; \ov{\PHa}^{\angles{k}}) = (\PHs; \PHl; \PHa^{\angles{2}})$
  the flattening of the PH with $k$ classes of priorities $\ov{\PHs}; \ov{\PHl}; \ov{\PHa}^{\angles{k}}$, with:
  \begin{itemize}
    \item $\PHs = \ov{\PHs} \cup \PHs_f$
      where $\PHs_f = \{ f^{h,i} \mid h \in \ov{\PHh} \wedge \n \geq 1 \wedge i \in \segm{1}{\n} \}$;
    \item $\PHl = \left( \underset{a \in \ov{\PHs}}{\times} \PHl_{a} \right) \times
      \left(\underset{f^{h,i} \in \PHs_f}{\times} \PHl_{f^{h,i}} \right)$,
      where $\forall a \in \ov{\PHs}, \PHl_{a} = \ov{\PHl}_{a}$, and
      $\forall f^{h,i} \in \PHs_f,
      \PHl_{f^{h,i}} = \{ f^{h,i}_\varsigma \mid \varsigma \in \underset{b \in \PHdep{i}{h}}{\times} \PHl_{b} \}$;
    \item $\PHh^{(1)} = \{ \PHhit{a_k}{f^{h,i}_\varsigma}{f^{h,i}_{\varsigma'}} \mid
      %h \in \ov{\PHh} \wedge %\n \geq 1 \wedge i \in \segm{1}{\n} \wedge
      f^{h,i} \in \PHs_f \wedge
      a \in \PHdep{i}{h} \wedge a_k \in \PHl_a \wedge
      f^{h,i}_\varsigma , f^{h,i}_{\varsigma'} \in \PHl_{f^{h,i}} \wedge
      \PHget{\varsigma}{a} \neq a_k \wedge \varsigma' = \varsigma \Cap \{ a_k \} \}$;
%      \todo{From here: ...} \PHget{\varsigma'}{a} = a_k \wedge
%      (\forall b \in \PHdep{i}{h} \setminus \{ a \}, \PHget{\varsigma'}{b} = \PHget{\varsigma}{b}) \}$;
%      \todo{... can be replaced by: $\varsigma' = \varsigma \Cap \{ a_k \}$}
    \item $\PHh^{(2)}=\{ \PHhit{f^{h,i}_\varsigma}{\target{h}}{\bounce{h}} \mid
      f^{h,i} \in \PHs_f \wedge
      %h \in \ov{\PHh} \wedge 
      f^{h,i}_\varsigma \in \PHl_{f^{h,i}} \wedge \Fsem{h}{\varsigma} \}$.
  \end{itemize}
  Given a state $s \in \PHl$, $\unflats{s} = \os$ is the corresponding state in $\ov{\PHl}$:
  $\forall a \in \ov{\PHs}, \PHget{\os}{a} = \PHget{s}{a}$.

  \noindent
  Given a state $\os \in \ov{\PHl}$, $\flats{\os} = s$ is the corresponding state in $\PHl$:
  $\forall a \in \ov{\PHs}, \PHget{s}{a} = \PHget{\os}{a}$
  and $\forall f^{h,i} \in \PHs_f, \PHget{s}{f^{h,i}} = f^{h,i}_\varsigma$ with $f^{h,i}_\varsigma \in \PHl_{f^{h,i}}$
  and $\forall b \in \PHdep{i}{h}, \PHget{\varsigma}{b} = \PHget{\os}{b}$.
\end{definition}

\begin{theorem}[$(\ov{\PHs}; \ov{\PHl}; \ov{\PHa}^{\angles{k}}) \approx \PHflat(\ov{\PHs}; \ov{\PHl}; \ov{\PHa}^{\angles{k}})$]
\label{thm:bisimPHP}
  Let $\ov{\PH} = (\ov{\PHs}; \ov{\PHl}; \ov{\PHa}^{\angles{k}})$ a PH with $k$ classes of priorities,
  and $\PH = \PHflat(\ov{\PHs}; \ov{\PHl}; \ov{\PHa}^{\angles{k}}) = (\PHs; \PHl; \PHa^{\angles{2}})$ its flattening.
  \begin{enumerate}
    \item \label{php2ph} $\forall \os, \os' \in \ov{\PHl}$,
      $\os \PHPtrans[\oPH] \os' \Longrightarrow \flats{\os} \PHPtrans[\oPH]^* \flats{\os'}$,
      where $\PHPtrans[\oPH]^*$ is a finite sequence of $\PHPtrans[\oPH]$ transitions.

    \item \label{ph2php} $\forall s, s' \in \PHl$,
      $s \PHPtrans s' \Longrightarrow \unflats{s} = \unflats{s'} \vee
      \unflats{s} \PHPtrans \unflats{s'}\enspace.$
  \end{enumerate}
\end{theorem}
%
\begin{proof}
  (\ref{php2ph}) Let $\flats{\os} = s$.
    From \pref{def:play}, if $\os \PHPtrans[\PH] \os'$,
    then there exists $h \in \ov{\PHh}$ so that $\os' = \os \PHplay h$.
    Furthermore, given \pref{th:ppplay}, $\Fsem{\F(h)}{\os}$.
    Finally, given \pref{def:flattening}, $\PHget{s}{f^{h,i}}$

  \todo{-- À continuer --}
\end{proof}

@@
